Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
df <- read.delim('pseudo_facebook.tsv')
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
#install.packages('ggthemes', dependencies = T)
#library(ggthemes)
qplot(x = dob_day, data = df) +
  scale_x_continuous(breaks=1:31)
```

***

This is the same as the code below:

```{r}
library(ggplot2)
ggplot(aes(x = dob_day), data = df) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

***
#### What are some things that you notice about this histogram?
Response: bod is highly concentrated on the 1st, while 31 has a low concentration. 31st will be normal, as there are no 31st in some months. The rest are near uniform.

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes: This works for each level of your categorical variable
facet_wrap(~VARIABLE)
facet_grid(VERTICAL~HORIZONTAL)

```{r Faceting}
ggplot(data = df, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month, ncol=4)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
ggplot(data = df, aes(x=friend_count)) +
  geom_histogram()
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
ggplot(data = df, aes(x=friend_count)) +
  geom_histogram() +
  scale_x_continuous(limits = c(0,1000))
```

### Exploring with Bin Width
Notes:
```{r}
ggplot(aes(x = friend_count), data = df) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
ggplot(aes(x = friend_count), data = df) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
ggplot(aes(x = friend_count), data = subset(df, !is.na(gender))) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
by(df$friend_count, df$gender, summary)
```

#### Who on average has more friends: men or women?
Response: women

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: since the distribution is long-tailed, the higher ends drag up the mean. Median is less sensitive to change, thus more robust.

***

### Tenure
Notes:

```{r Tenure}
ggplot(aes(x = tenure), data = df) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
ggplot(aes(x=tenure/365), data = df) +
  geom_histogram(binwidth = 25/365, color = 'black', fill = '#F79420')
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
ggplot(aes(x=tenure/365), data = df) +
  geom_histogram(binwidth = 25/365, color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0,7)) +
  xlab('# years using Facebook') + 
  ylab('# users in sample')
```

***

### User Ages
Notes:

```{r User Ages}
ggplot(aes(x=age), data = df) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 115, 5), limits = c(0,115))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
```{r}
summary(df$friend_count)
#this has -inf, as some has friend_count == 0.
summary(log10(df$friend_count))
#counter by adding 1 overall
summary(log10(df$friend_count + 1))

#3 plots
install.packages('gridExtra')
library(gridExtra)

p1 = ggplot(aes(x = friend_count), data = df) +
  geom_histogram(binwidth = .25)
p2 = p1 + scale_x_log10()
p3 = p1 + scale_x_sqrt()
grid.arrange(p1, p2, p3)
```

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
logScale <- ggplot(aes(x = log10(friend_count)), data = df) +
  geom_histogram()

#preferred
countScale <- ggplot(aes(x = friend_count), data = df) +
  geom_histogram() +
  scale_x_log10()

grid.arrange(logScale, countScale, ncol = 2)
```

***


### Frequency Polygons

```{r Frequency Polygons}
ggplot(aes(x=friend_count, y=..count../sum(..count..)), 
       data = subset(df, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth = 10) +
  #scale_x_continuous(limits = c(0,1000), breaks = seq(0, 1000, 50))
  scale_x_log10()

ggplot(aes(x = www_likes), data = subset(df, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10()
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
by(df$www_likes, df$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
ggplot(aes(x = gender, y = friend_count), data = subset(df, !is.na(gender))) +
  geom_boxplot()
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
ggplot(aes(x = gender, y = friend_count), data = subset(df, !is.na(gender))) +
  geom_boxplot() + 
  coord_cartesian(ylim = c(0, 1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
ggplot(aes(x = gender, y = friend_count), data = subset(df, !is.na(gender))) +
  geom_boxplot() + 
  coord_cartesian(ylim = c(0, 250))

by(df$friend_count, df$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
ggplot(aes(x = gender, y = friendships_initiated), data = subset(df, !is.na(gender))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0,150))
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
mobile_check_in <- NA
df$mobile_check_in <- ifelse(df$mobile_likes > 0, 1, 0)
df$mobile_check_in <- factor(df$mobile_check_in)
summary(df$mobile_check_in)[2] / (summary(df$mobile_check_in)[1] + summary(df$mobile_check_in)[2])
#or
sum(df$mobile_check_in == 1)/length(df$mobile_check_in)
```

Response:

***

### Analyzing One Variable
Reflection:
Close look (zooming in) important. Types, distribution, missing, outliers. W. histogram, freqpoly, boxplot.
Adjustments - limits, log / sqrt , binaries...

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!