Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
df <- read.delim('pseudo_facebook.tsv')
ggplot(aes(x = age, y = friend_count), data = df) +
  geom_point(alpha = 1/20) +
  xlim(13, 90)
summary(df$age)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = df) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = df) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = df) +
  geom_point(alpha = 1/20) +
  xlim(13, 90) +
  coord_trans(y = "sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x=age, y=friendships_initiated), data = df) +
  geom_point(alpha=1/20, position = position_jitter(h = 0))  +
  coord_trans(y = "sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

age_groups <- group_by(df, age)
df.fc_by_age <- summarise(age_groups,
          friendcount_mean = mean(friend_count),
          friendcount_median = median(friend_count),
          n = n())
df.fc_by_age <- arrange(df.fc_by_age, age)

#same as...
df.fc_by_age <- df %>%
  group_by(age) %>%
  summarise(friendcount_mean = mean(friend_count),
          friendcount_median = median(friend_count),
          n = n()) %>%
  arrange(age)

```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x=age, y=friendcount_mean), data = df.fc_by_age) +
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age, y=friendships_initiated), data = df) +
  geom_point(alpha=1/20, position = position_jitter(h = 0), color = "orange")  +
  coord_trans(y = "sqrt") +
  geom_line(stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .1), linetype = 2, color = "blue") +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = .9), linetype = 2, color = "blue") +
  coord_cartesian(xlim = c(13, 30), ylim = c(0, 2000))
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(df$age, df$friend_count, method = "pearson")
#alternative
with(df, cor.test(age, friend_count, method = "pearson"))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(df, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x=www_likes_received, y=likes_received), data = df) +
  geom_point(alpha = 1/20) +
  coord_cartesian(xlim = c(0, 2000), ylim = c(0,2000))
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x=www_likes_received, y=likes_received), data = df) +
  geom_point(alpha = 1/20) +
  coord_cartesian(xlim = c(0, quantile(df$www_likes_received, 0.95)), ylim = c(0,quantile(df$likes_received, 0.95)))+
  geom_smooth(method = "lm", color = "red")
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(df, cor.test(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
dim(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Month, Temp))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 203, 12))

#or modulo:
ggplot(aes(x = (Month%%12), y=Temp), data = Mitchell) +
  geom_point()
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x=age, y=friendcount_mean), data = df.fc_by_age) +
  geom_line()
df$age_with_months <- df$age + (1 - df$dob_month / 12)
```

***

### Age with Months Means

```{r Age with Months Means}
df.fc_by_age_months <- df %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)
```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y= friend_count_mean), data = subset(df.fc_by_age_months, age_with_months < 71)) +
  geom_line()
```

***

### Smoothing Conditional Means
Notes: By increasing the size of the bins, the noise is smoothed, as a bunch. 
Less data points, wider binwidth -> more precise means, but may miss out on the important features of the relationship.

"Bias - Variance Tradeoff"

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age_with_months, y= friend_count_mean), 
             data = subset(df.fc_by_age_months, age_with_months < 71)) +
  geom_line() +
  geom_smooth()

p2 <- ggplot(aes(x=age, y=friendcount_mean), data = subset(df.fc_by_age, age < 71)) +
  geom_line() +
  geom_smooth()

p3 <- ggplot(aes(x=round(age/5) * 5, y=friend_count), data = subset(df, age < 71)) +
  geom_line(stat = "summary", fun.y = mean)

library(gridExtra)
grid.arrange(p1, p2, p3)
```

***

### Which Plot to Choose?
Notes: You don't have to choose. We can have multiple viz & summaries for the same data.
geom_smooth() can help to see a smoothed trend in the relationship. However, still subject to systematic errors (for example, discontinuity on age 69)

***

### Analyzing Two Variables
Reflection:
Scatter plot the main, but also with conditional summaries, such as means. Also, alpha / jitter. 
Do not trust the first data that you see.
Correlation lets you decide which data to use.
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

