Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes: Tried color, did not tell much

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
df <- read.delim('pseudo_facebook.tsv')
df.fc_by_age_gender <- subset(df, !is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) 
head(df.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count), data = df.fc_by_age_gender) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
df.fc_by_age_gender.wide <- dcast(df.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = "median_friend_count")
head(df.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
#df.fc_by_age_gender.wide$ratio = df.fc_by_age_gender.wide$female / df.fc_by_age_gender.wide$male
#ggplot(aes(x = age, y = ratio), data = df.fc_by_age_gender.wide) +
ggplot(aes(x = age, y = female/male), data = df.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = 2, color = "blue", alpha = 0.2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
df$year_joined <- floor(2014 - (df$tenure / 365))
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
df$year_joined.bucket <- cut(df$year_joined, c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count), data = subset(df, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count), data = subset(df, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(df, tenure > 0), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate? 
0.2205
What is the maximum friend rate?
417.0000
```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure), data = subset(df, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

#using geom smooth
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(df, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes: used positive/negative word ratio. Not enough, needed to model  -> used moving averages model
Was too noisy, so aggregated to one-day moving average.
Predicting model -> needed to get rid of natural splines -> summarized way too much.
Then aggregated it up for 7days moving averages, and was able to see the pattern.
For simple moving averages, very flexible. More lags to moving average -> lower variance, but high bias
Used splines to accomodate

***

### Introducing the Yogurt Data Set
Notes:
```{r}
yo <- read.csv("yogurt.csv")
str(yo)

yo$id <- factor(yo$id)
str(yo)
```

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
p1 <- ggplot(aes(x=price), data = yo) +
  geom_histogram()

p2 <- ggplot(aes(x=price), data = yo) +
  geom_histogram(binwidth = 10)

library(gridExtra)
grid.arrange(p1, p2, ncol=2)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price)
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)

#or
yo$all.purchases <- yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry
```

***

### Prices over Time
Notes: Common Price do rise over time

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yo) +
  geom_point(alpha = .05, position = "jitter")
```

***

### Sampling Observations
Notes:
Multiple observation of same units -> sampling is more helpful as it visualizes well.
***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(630)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) +
  facet_wrap( ~ id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)
```

People tend to buy more when the price is lower. 
***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
#install.packages('GGally')
library(GGally)

set.seed(1836)
df_subset <- df[, c(2:15)]
names(df_subset)
ggpairs(df_subset[sample.int(nrow(df_subset), 1000), ])
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

